---
title: "ESM 204 Assignment 4"
author: "Sam Rozal"
date: "5/16/2022"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
```


```{r}
#read in the data 
warming <- read_csv(here("data","warming.csv")) 

damages <- read_csv(here("data","damages.csv"))
```
# Question 1 

```{r}
#create a model for climate change
climate_change_damages <- lm (damages ~ warming + I(warming^2) - 1, data=damages)

#get the coefficients from the model, so later on we can write a function with the coefficients for predicting stuff later
summary(climate_change_damages)

#get predicted values 
damages_predicted <- climate_change_damages$fitted.values
```
The equation of the line is damages = -1.06e12(warming)+ 1.941e13(warming^2)

```{r}
#write a function to display this equation

damage_function <- function(warming)
  {dam <- (climate_change_damages$coefficients[1]* (warming)) +
    (climate_change_damages$coefficients[2] * (warming^2))
  return(dam)}
```

```{r}
ggplot(data=climate_change_damages, aes(x=warming,y=damages))+
  geom_line(aes(x=warming,y=damages_predicted), size = 1.15, color = "royalblue")+
  geom_point(aes(x=warming,y=damages), shape = 21, color = "black", fill = "grey")+
  labs(x="Warming (Celsius)", y="Damages (USD)", title="Damages from Increased Global Mean Temperature")+
  theme_minimal()+
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))+
  theme(legend.background= element_rect(fill= "azure1"))
```


# Question 2
Warming < baseline , pulse 
predict using Question 1 model 
baseline over years 
- Step 1 create a new tibble 
baseline = tibble(warming = warming$warming_baseline)
predicted_baseline = predict(model from question 1, newdata= baseline)

2. Use warming.csv and your estimated damage function to predict damages in each year under
the baseline climate and the pulse scenario. Make four plots: (1) damages over time without
the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that
arises from the pulse, and (4) the difference in damages over time from the pulse per ton of
CO2 (you can assume that each ton of the pulse causes the same amount of damage).

### Plot 1 - damages over time without pulse
```{r}
#create a new tibble to get the values of predicted baseline
baseline = tibble(warming = warming$warming_baseline)
predicted_baseline = predict(climate_change_damages, newdata= baseline)
```

```{r}
#use the written function to get the values for the dmage at the baseline 
damages_baseline_df <- warming %>%
  mutate(baseline_damages = damage_function(warming_baseline))
```

```{r}
#graph it 

ggplot(data=damages_baseline_df, aes(x=year, y=baseline_damages))+
  geom_line(size=1.3, color = "royalblue")+
  labs(title="Baseline Damages", x="Year", y ="Damages (USD)")+
  theme_minimal()+
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))+
  theme(legend.background= element_rect(fill= "azure1"))
```
### Plot  - damages over time with pulse
```{r}
#create a new data frame with damages with damages with and without pulse
damages_baseline_pulse_df <- warming %>%
  mutate(baseline_damages = damage_function(warming_baseline))%>%
  mutate(pulse_damages = damage_function(warming_pulse))
```

```{r}
ggplot(data=damages_baseline_pulse_df)+
  geom_line(aes(x=year, y=pulse_damages), size = 1.3, color = "royalblue")+
  labs(title="Damages with 35 tons of CO2 pulse", x="Year", y ="Damages (USD)")+
  theme_minimal()+
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))+
  theme(legend.background= element_rect(fill= "azure1"))
```
### Plot 3 -the difference in damages over time that arises from the pulse
```{r}
#create a new data frame with damages with damages with pulse, without pulse, and difference in damages
damages_baseline_pulse_diff_df <- warming %>%
  mutate(baseline_damages = damage_function(warming_baseline))%>%
  mutate(pulse_damages = damage_function(warming_pulse)) %>% 
  mutate(diff_damages = (pulse_damages-baseline_damages))
```

```{r}
ggplot(data=damages_baseline_pulse_diff_df)+
  geom_line(aes(x=year, y=diff_damages), size = 1.3, color = "royalblue")+
  labs(title="Difference in Damages With and Without Pulse", x="Year", y ="Damages(USD)")+
  theme_minimal()+
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))+
  theme(legend.background= element_rect(fill= "azure1"))
```

### Plot 4 - the difference in damages over time from the pulse per ton of CO2 (you can assume that each ton of the pulse causes the same amount of damage)

```{r}
#create a new data frame with damages with damages with pulse, without pulse, difference in damages, and differnce in damages per ton Co2
damages_baseline_pulse_diff_perton_df <- warming %>%
  mutate(baseline_damages = damage_function(warming_baseline))%>%
  mutate(pulse_damages = damage_function(warming_pulse)) %>% 
  mutate(diff_damages = (pulse_damages-baseline_damages)) %>% 
  mutate(diff_damages_perton = (diff_damages/3500000000000))
```

```{r}
ggplot(data=damages_baseline_pulse_diff_perton_df)+
  geom_line(aes(x=year, y=diff_damages_perton), size = 1.3, color = "royalblue")+
  labs(title="Difference in Damages (With and Without Pulse) Per Ton", x="Year", y ="Damages(USD)")+
  theme_minimal()+
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))+
  theme(legend.background= element_rect(fill= "azure1"))
```

# Question 3
3. The SCC is the present discounted value of the stream of future damages caused by one additional ton of CO2. The Obama Administration used a discount rate of 3% to discount damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates. Explain the intuition for how the discount rate affects the SCC.

Note: The equation to calculate present value is PV= B/((1+r^t))

- We assume that a reasonable range of discount rates is 1%-10%. This is because _____________.

```{r}
# Write a function for present value 
PV <- function(damages_per_ton,discount_rate, year, pulse_year){
  pv= (damages_per_ton)/ (((1+discount_rate)^year-pulse_year))
return(pv)
}
```

```{r}
PV(damages_baseline_pulse_diff_perton_df$diff_damages_perton,0.1,2023,2022)
```

```{r}
# Set the pulse year to 2022
pulse_year <- 2022

discount_rates<- damages_baseline_pulse_diff_perton_df %>% 
  mutate(discount_rate_1_percent = diff_damages_perton / ((1.01)^(year-pulse_year))) %>% 
  mutate(discount_rate_2_percent = diff_damages_perton / ((1.02)^(year-pulse_year))) %>% 
  mutate(discount_rate_3_percent = diff_damages_perton / ((1.03)^(year-pulse_year))) %>% 
  mutate(discount_rate_4_percent = diff_damages_perton / ((1.04)^(year-pulse_year))) %>% 
  mutate(discount_rate_5_percent = diff_damages_perton / ((1.05)^(year-pulse_year))) %>% 
  mutate(discount_rate_6_percent = diff_damages_perton / ((1.06)^(year-pulse_year))) %>% 
  mutate(discount_rate_7_percent = diff_damages_perton / ((1.07)^(year-pulse_year))) %>% 
  mutate(discount_rate_8_percent = diff_damages_perton / ((1.08)^(year-pulse_year))) %>% 
  mutate(discount_rate_9_percent = diff_damages_perton / ((1.09)^(year-pulse_year))) %>% 
  mutate(discount_rate_10_percent = diff_damages_perton / ((1.1)^(year-pulse_year))) 

discount_rates_simplified <- discount_rates %>% 
  select(year,discount_rate_1_percent,discount_rate_2_percent,discount_rate_3_percent,discount_rate_4_percent,discount_rate_5_percent,discount_rate_6_percent,discount_rate_7_percent,discount_rate_8_percent,discount_rate_9_percent,discount_rate_10_percent)
```

```{r}
#create a data frame with just the sums of the present values under each of the discount rates
social_cost_df <- data.frame((present_values <- 
                                c(sum(discount_rates_simplified$discount_rate_1_percent),
                              sum(discount_rates_simplified$discount_rate_2_percent),
                              sum(discount_rates_simplified$discount_rate_3_percent),
                              sum(discount_rates_simplified$discount_rate_4_percent),
                              sum(discount_rates_simplified$discount_rate_5_percent),
                              sum(discount_rates_simplified$discount_rate_6_percent),
                              sum(discount_rates_simplified$discount_rate_7_percent),
                              sum(discount_rates_simplified$discount_rate_8_percent),
                              sum(discount_rates_simplified$discount_rate_9_percent),
                              sum(discount_rates_simplified$discount_rate_10_percent))))

#create a data frame with just the dicount rates
discount_rates <- data.frame((discount_rate <- 
                                c(0.01,0.02,0.03,0.04,0.05,0.06,0.07,0.08,0.09,0.1)))

#join the data frames for funsies

                             
```


```{r}
#create a vector with the differnet presnt values
present_values <- c(sum(discount_rates_simplified$discount_rate_1_percent),
                    sum(discount_rates_simplified$discount_rate_2_percent),
                    sum(discount_rates_simplified$discount_rate_3_percent),
                    sum(discount_rates_simplified$discount_rate_4_percent),
                    sum(discount_rates_simplified$discount_rate_5_percent),
                    sum(discount_rates_simplified$discount_rate_6_percent),
                    sum(discount_rates_simplified$discount_rate_7_percent),
                    sum(discount_rates_simplified$discount_rate_8_percent),
                    sum(discount_rates_simplified$discount_rate_9_percent),
                    sum(discount_rates_simplified$discount_rate_10_percent))

#make it inti a data frame
social_cost_df <- data.frame(present_values)

#create a vector of the diffenrt discount rates
discount_rate <- c(0.01,0.02,0.03,0.04,0.05,0.06,0.07,0.08,0.09,0.1)

#make it into a data frame
discount_rate_df <- data.frame(discount_rate) 

#make a data frame with both discont rates and present values
scc_and_discount_df <- data.frame(present_values,discount_rate)

```

```{r}
#create a graph with the discount rates and present values 

ggplot(data=scc_and_discount_df, aes(x=discount_rate, y=present_values))+
  geom_line()+
  labs(title = "Social Cost of Carbon at Different Discount Rates", y="Social Cost of Carbon (Present Value)", x= "Discount Rate")+
  theme_minimal()+
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))+
  theme(legend.background= element_rect(fill= "azure1"))

```

# Question 4 
```{r}

```

