---
title: "ESM 204 Assignment 4"
author: "Sam Rozal"
date: "5/16/2022"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
```


```{r}
#read in the data 
warming <- read_csv(here("data","warming.csv")) 

damages <- read_csv(here("data","damages.csv"))
```
# Question 1 

```{r}
#create a model for climate change
climate_change_damages <- lm (damages ~ warming + I(warming^2) - 1, data=damages)

#get the coefficients from the model, so later on we can write a function with the coefficients for predicting stuff later
summary(climate_change_damages)

#get predicted values 
damages_predicted <- climate_change_damages$fitted.values
```
The equation of the line is damages = -1.06e12(warming)+ 1.941e13(warming^2)

```{r}
#write a function to display this equation

damage_function <- function(warming)
  {dam <- climate_change_damages$coefficients[1]* (warming) +
    climate_change_damages$coefficients[2] * (warming^2)
  return(dam)}
```

```{r}
ggplot(data=climate_change_damages, aes(x=warming,y=damages))+
  geom_line(aes(x=warming,y=damages_predicted))+
  geom_point(aes(x=warming,y=damages))+
  labs(x="Warming (Celsius)", y="Damages from Warming(USD)", title="Damages (in USD) from Increased Global Mean Temperture")+
  theme_classic()+
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))+
  theme(legend.background= element_rect(fill= "azure1"))
```


# Question 2
Warming < baseline , pulse 
predict using Question 1 model 
baseline over years 
- Step 1 create a new tibble 
baseline = tibble(warming = warming$warming_baseline)
predicted_baseline = predict(model from question 1, newdata= baseline)

2. Use warming.csv and your estimated damage function to predict damages in each year under
the baseline climate and the pulse scenario. Make four plots: (1) damages over time without
the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that
arises from the pulse, and (4) the difference in damages over time from the pulse per ton of
CO2 (you can assume that each ton of the pulse causes the same amount of damage).

### Plot 1 - damages over time without pulse
```{r}
#create a new tibble to get the values of predicted baseline
baseline = tibble(warming = warming$warming_baseline)
predicted_baseline = predict(climate_change_damages, newdata= baseline)
```

```{r}
#use the written function to get the values for the dmage at the baseline 
damages_baseline_df <- warming %>%
  mutate(baseline_damages = damage_function(warming_baseline))
```

```{r}
#graph it 

ggplot(data=damages_baseline_df, aes(x=year, y=baseline_damages))+
  geom_line()+
  labs(title="Baseline Damages", x="Year", y ="Damages (USD)")+
  theme_classic()+
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))+
  theme(legend.background= element_rect(fill= "azure1"))
```
### Plot  - damages over time with pulse
```{r}
#create a new data frame with damages with damages with and without pulse
damages_baseline_pulse_df <- warming %>%
  mutate(baseline_damages = damage_function(warming_baseline))%>%
  mutate(pulse_damages = damage_function(warming_pulse))
```

```{r}
ggplot(data=damages_baseline_pulse_df)+
  geom_line(aes(x=year, y=pulse_damages))+
  labs(title="Damages wih 35 tons of CO2 pulse", x="Year", y ="Damages (USD)")+
  theme_classic()+
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))+
  theme(legend.background= element_rect(fill= "azure1"))
```
### Plot 3 -the difference in damages over time that arises from the pulse
```{r}
#create a new data frame with damages with damages with pulse, without pulse, and difference in damages
damages_baseline_pulse_diff_df <- warming %>%
  mutate(baseline_damages = damage_function(warming_baseline))%>%
  mutate(pulse_damages = damage_function(warming_pulse)) %>% 
  mutate(diff_damages = (warming_pulse-warming_baseline))
```

```{r}
ggplot(data=damages_baseline_pulse_diff_df)+
  geom_line(aes(x=year, y=diff_damages))+
  labs(title="Differnece in Damages with and without pulse", x="Year", y ="Damages(USD)")+
  theme_classic()+
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))+
  theme(legend.background= element_rect(fill= "azure1"))
```
### Plot 4 - 